Three Algorithms for Merging Navigable Graphs

### 2024-10-08 paper/main.tex
finish pseudocode for merge2 algorithm
add  pseudocode for merge1 algorithm

\usepackage{natbib}
\usepackage{graphicx}
\usepackage{algorithm}
\usepackage{algorithmicx}
\usepackage[noend]{algpseudocode}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\begin{document}

@@

\section {Local}

@@

\State $v^*$ \gets $v_0$ 
\For{$i = l_{max} \; \textbf{to} \; \ell $}
    \State $v^*$ \gets \textsc{LocalSearch}($G=G_i, q=q, C={v^*}, k=1, L=L$)
\EndFor

@@

\section{Merge2}

\begin{algorithm}
\caption{\textsc{HNSW-Merge-Basic}($\mathcal{H}_a, \mathcal{H}_b$)}\label{alg:general_merge}
\textbf{Require:} The HNSW graphs $\mathcal{H}_a = (G^a_i), \mathcal{H}_b = (G^b_i)$ \\
\textbf{Ensure:}  Merged HNSW graphs $\mathcal{H}_c = (G^c_i)$ 
\begin{algorithmic}[1]

@@

\begin{algorithm}
\caption{\textsc{Merge2}($G^a, G^b, \ell, \text{jump\_ef}, \text{local\_ef}, \text{next\_step\_k}, M$)}\label{alg:merge2}
\textbf{Require:} The graphs $G^a=(V^a,E^a), G^b=(V^b,E^b)$ \\
\textbf{Ensure:}  Merged HNSW graphs $\mathcal{H}_c = (G^c_i)$ 
\begin{algorithmic}[1]

@@

    \State $\mathcal{P}_a  \gets \textsc{HNSW-Search}(\mathcal{H}=\mathcal{H}^a, q=v*, v_0, k, L=\text{jump\_ef}, \ell $)

    \State $\mathcal{P}_b \gets \textsc{HNSW-Search}(\mathcal{H}=\mathcal{H}^b, q=v*, v_0, k, L=\text{jump\_ef}, \ell $)

@@

        \State $\mathcal{C}^a \gets \text{local search in } \mathcal{H}_a.\mathcal{G}[\ell] \text{ from } D_{merged}[v^*]$
        \State $\mathcal{C}^b \gets \text{local search in } \mathcal{H}_b.\mathcal{G}[\ell] \text{ from } D_{merged}[v^*]$

        \If{$v^* \in V^a $}
            \State $\mathcal{C}_{merged} \gets  \{v : (v^*, v) \in E^a \} \cup  \mathcal{C}^b\}$
        \Else
            \State $\mathcal{C}_{merged} \gets  \{v : (v^*, v) \in E^b \} \cup  \mathcal{C}^a\}$
        \EndIf

@@

        \State $E^c \gets E^c \cup \{ (v^*, v) : v \in \text{neighborhood\_construction}(\mathcal{C}_{merged}, v^*, m) \}$

        \State $\mathcal{C}^a_{not\_done} \gets \{(v \in \{\mathcal{C}^a_1, \mathcal{C}^a_2, ..., \mathcal{C}^a_{ \text{next\_step\_k} } \} \mid v \in \mathcal{V}_{not\_done} \}$

        \State $\mathcal{C}^b_{not\_done} \gets \{(v \in \{\mathcal{C}^b_1, \mathcal{C}^b_2, ..., \mathcal{C}^b_{ \text{next\_step\_k} } \} \mid v \in \mathcal{V}_{not\_done} \}$

@@

        \State $\text{new\_curr} \gets \arg\min_{(i, d) \in \mathcal{C}^{not\_done}} d$
        \State $v^* \gets \text{new\_curr}$
        \State $\mathcal{P}_a \gets \{i \mid (i, d) \in \mathcal{C}_a\}$
        \State $\mathcal{P}_b \gets \{i \mid (i, d) \in \mathcal{C}_b\}$
    \EndWhile
\EndWhile

@@

>>>>

\usepackage{natbib}
\usepackage{graphicx}
\usepackage{algorithm}
\usepackage{algorithmicx}
\usepackage[noend]{algpseudocode}
\usepackage{amsmath}
\DeclareMathOperator*{\argminA}{arg\,min} % Jan Hlavacek
\DeclareMathOperator*{\argminB}{argmin}   % Jan Hlavacek
\usepackage{amsfonts}
\usepackage{amssymb}
\begin{document}

@@

\section {Search}

@@

\State $v^* \gets v_0$ 
\For{$i = l_{max} \; \textbf{to} \; \ell $}
    \State $v^* \gets \textsc{LocalSearch}(G=G_i, q=q, C={v^*}, k=1, L=L)$
\EndFor

@@

\section {Neighborhood Construction Strategies}

There are different strategies to form a vertex neighborhood. Our proposed merge algorithms are independent in neighborhood procedures. Merge algorithm can use a neighborhood construction procedure as an input parameter.

\begin{algorithm}
\caption{\textsc{KNN-Neighborhood-Construction}$(u,U, k)$}
\label{alg:knntrategy}
\textbf{Require:} vertex $u \in V$, neighbor candidates $U \subseteq V$, $k \in \mathbb{N} $\\
\textbf{Ensure:} selected neighbors $U' \subseteq U$
\begin{algorithmic}[1]

    \State $U' \gets \text{k-closest neighbors of } U \text{ to } u$
    \State \Return $U'$
\end{algorithmic}
\end{algorithm}

\begin{algorithm}
\caption{\textsc{RNG-Neighborhood-Construction}$(u, U, m)$}
\label{alg:rngstrategy}
\textbf{Require:} vertex $u \in V$, neighbor candidates $U \subseteq V$, maximum neighborhood size $m \in \mathbb{N}$\\
\textbf{Ensure:} selected neighbors $U' \subseteq U$
\begin{algorithmic}[1]
    \State sort $v \in U$ in ascending order of $\delta(u,v)$
    \State $U' \gets \emptyset$
    \For{$v \in U$}
        \State $f \gets \text{true}$
        \For{$w \in U'$}
            \If{$\delta(u,v) \geq \delta(v,w)$}
                \State $f \gets \text{false}$
                \State \textbf{break}
            \EndIf
        \EndFor
        \If{$f$}
            \State $U' \gets U' \cup \{v\}$
        \EndIf

        \If{$|U'| \geq m $}
            \State \textbf{break}
        \EndIf
    \EndFor
    \State \Return $U'$
\end{algorithmic}
\end{algorithm}

\section{Merge Algorithms}

\subsection{HNSW Merge Basic}

\begin{algorithm}
\caption{\textsc{HNSW-General-Merge}($\mathcal{H}_a, \mathcal{H}_b$)}\label{alg:general_merge}
\textbf{Require:} The HNSW graphs $\mathcal{H}_a = (G^a_i), \mathcal{H}_b = (G^b_i)$ \\
\textbf{Ensure:}  Merged HNSW graphs $\mathcal{H}_c = (G^c_i)$ 
\begin{algorithmic}[1]

@@

\subsection{Merge1}

\begin{algorithm}
\caption{\textsc{Merge1}($\mathcal{H}_a, \mathcal{H}_b, \ell, \text{jump\_ef}, \text{local\_ef}, \text{next\_step\_k}, M, m$)}\label{alg:merge1}
\textbf{Require:} The HNSW graphs $\mathcal{H}_a = (G^a_i), \mathcal{H}_b = (G^b_i)$, the size of the forming neighborhoods $m \in \mathbb{N} $ \\
\textbf{Ensure:}  Merged HNSW graphs $\mathcal{H}_c = (G^c_i)$ 
\begin{algorithmic}[1]

\State $E^c \gets \emptyset$
\State $\mathcal{V}_{not\_done} \gets V^a$

    \State $\mathcal{P}^b  \gets \textsc{HNSW-Search}(\mathcal{H}=\mathcal{H}^b, q=v^*, v_0, k=M, L=\text{jump\_ef}, \ell)$

        \State $\mathcal{C}^b  \gets \textsc{LocalSearch}(G=G^b, q=v^*, C=\mathcal{P}^b , k=m, L=\text{local\_ef})$

        \State $\mathcal{C} \gets  \{v : (v^*, v) \in E^a \} \cup \mathcal{C}^b$

        \State $E^c[v^*] \gets \textsc{NeighborhoodConstruction}(\mathcal{C}, v^*, m, D_{merged})$

        \State $\mathcal{C}_{not\_done} \gets \{\mathcal{C}^b_1, \mathcal{C}^b_2, ..., \mathcal{C}^b_{ \text{next\_step\_k} } \} \cap \mathcal{V}_{not\_done}$

        \State $v^* \gets \underset{v \in \mathcal{C}_{not\_done}}{\mathrm{argmin}} \rho(v, v^*)$
    \EndWhile
\EndWhile

\State $\mathcal{V}_{not\_done} \gets V^b$

\While{$\mathcal{V}_{not\_done} \neq \emptyset$}
    \State Repeat the same process for $V^b$ with the roles of $\mathcal{H}_a$ and $\mathcal{H}_b$ swapped.
\EndWhile

\subsection{Merge2}

\begin{algorithm}
\caption{\textsc{Merge2}($G^a, G^b, \ell, \text{jump\_ef}, \text{local\_ef}, \text{next\_step\_k}, M, m$)}\label{alg:merge2}
\textbf{Require:} The graphs $G^a=(V^a,E^a), G^b=(V^b,E^b)$, the size of the forming neighborhoods $m \in \mathbb{N} $ \\
\textbf{Ensure:}  Merged HNSW graphs $\mathcal{H}_c = (G^c_i)$ 
\begin{algorithmic}[1]

@@

    \State $\mathcal{P}^a  \gets \textsc{HNSW-Search}(\mathcal{H}=\mathcal{H}^a, q=v^*, v_0, k, L=\text{jump\_ef}, \ell $)

    \State $\mathcal{P}^b \gets \textsc{HNSW-Search}(\mathcal{H}=\mathcal{H}^b, q=v^*, v_0, k, L=\text{jump\_ef}, \ell $)

@@

        \State $ \mathcal{C}^a  \gets \textsc{LocalSearch}(G=G^a, q=v^*, C=\mathcal{P}^a  , k=m, L=\text{local\_ef})$

        \State $\mathcal{C}^b \gets \textsc{LocalSearch}(G=G^b, q=v^*, C=\mathcal{P}^b , k=m, L=\text{local\_ef})$

        \If{$v^* \in V^a $}
            \State $\mathcal{C}\gets  \{v : (v^*, v) \in E^a \} \cup  \mathcal{C}^b\}$
        \Else
            \State $\mathcal{C} \gets  \{v : (v^*, v) \in E^b \} \cup  \mathcal{C}^a\}$
        \EndIf

@@

        \State $E^c \gets E^c \cup \{ (v^*, v) : v \in \text{neighborhood\_construction}(\mathcal{C}, v^*, m) \}$

        \State $\mathcal{C}^a_{not\_done} \gets \{\mathcal{C}^a_1, \mathcal{C}^a_2, ..., \mathcal{C}^a_{ \text{next\_step\_k} } \} \cap \mathcal{V}_{not\_done}$

        \State $\mathcal{C}^b_{not\_done} \gets \{\mathcal{C}^b_1, \mathcal{C}^b_2, ..., \mathcal{C}^b_{ \text{next\_step\_k} } \}  \cap \mathcal{V}_{not\_done}$

@@

        \State $v^* \gets \underset{v \in \mathcal{C}_{not\_done}}{\mathrm{argmin}} \rho(v, v^*)$

        \State $\mathcal{P}_a \gets \mathcal{C}^a$
        \State $\mathcal{P}_b \gets \mathcal{C}^b$
    \EndWhile
\EndWhile

@@

\section{Computational Experiments}
TODO
